---
title: "Analysis of Data generate on Single-node Spark Cluster"
author: "Bernhard Bielesch"
output: 
  html_document: 
    highlight: tango
    theme: cerulean
---
```{r setup}
rm(list = ls()) # clean global environment
includeDir <- "../include"
source(file.path(includeDir, "includeRmd.R"))

platform <- "Local" # Platfrom which generated data to analyze, either "Local" or "Scape01"
```

## Load necessary data
```{r}
#####################################################################

load(file = file.path(dataDir, paste0("outputParamsExt-", platform, ".Rda")))
load(file = file.path(dataDir, paste0("applicationsRuntimeExt-", platform, ".Rda")))
load(file = file.path(dataDir, paste0("jobsRuntime-", platform, ".Rda")))
```

## Build Random Forest to verify Variable Importance for Accuracy

### Preparations for Forest Building
```{r}
dropCols <- c("id", "appName", "cluster",
              "startTime", "endTime",
              "lastUpdated", "features",
              "featuresName", "split",
              "testObservations", "duration", "testError")
factorCols <- c("algorithm", "classification")
integerCols <- c("dimensions", "trainObservations")

treeAppAccuracy <- 
  applicationsRuntimeExt %>%
  mutate_at(.vars = factorCols, .fun = funs(factor(.))) %>%
  mutate_at(.vars = integerCols, .fun = funs(as.integer(.))) %>% 
  mutate(dimensions = as.integer(dimensions)) %>%
  mutate(trainObservations = as.integer(trainObservations)) %>%
  mutate(accuracy = 1 - testError) %>%
  select(-one_of(dropCols))

Desc(treeAppAccuracy)

```

### Building the forest and visualize Variable Importance
```{r}
#####################################################################
## Random Forest overall
#####################################################################

# prepare tree building
response <- "accuracy"
features <- setdiff(names(treeAppAccuracy), c(response))
mlformula <- reformulate(termlabels = features, response = response)  

set.seed(50040)
rfModel <- randomForest(mlformula, data = treeAppAccuracy, 
                        ntree = 150, keep.inbag = TRUE, importance = TRUE)
```

### Visualize Variable Importance
```{r}
print(rfModel)
importance(rfModel)
```

```{r}
varImpPlot(rfModel, main = "Variable Importance for Accuracy")
```

### Visualize Variable Importance (Alternative)
```{r}
# Extracts variable importance and processed information
varImportance <- data.frame(importance(rfModel)) 
varImportance %<>%
  mutate(variable = as.factor(rownames(.))) %>%
  rename(PercIncMSE = X.IncMSE) %>%
  arrange(desc(PercIncMSE))
               
varImportance %>%
  ggplot(aes(x = variable, weight = PercIncMSE, fill = variable)) +
  geom_bar() + ggtitle("Variable Importance from Random Forest for Accuracy") + 
  labs(x = "Variable", y = "Variable Importance (Mean Decrease in MSE)") +
  scale_x_discrete(limits = as.character(varImportance$variable)) +
  scale_fill_discrete(name = "Variable Name",
                      breaks = as.character(varImportance$variable))
# ggsave(file = file.path(plotDir, platform,
#                         paste0("accuracyVariableImportance-", platform, ".png")))
```

## Visualization of Accuracy/Error rate
```{r}
applicationsRuntimeExt %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  mutate(accuracy = 1 - testError) %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = trainObservations, color = accuracy, 
                            size = accuracy, shape = algorithm), alpha = 0.5) +
  labs(color = "Accuracy", size = "Accuracy") +
  ggtitle("Accuracy for dimensions/training observations") +
  labs(x = "Dimensions", y = "No of Training Observations") +
  scale_color_gradient(labels = comma, low = "red", high = "green") +
  scale_y_continuous(labels = comma)
# ggsave(file = file.path(plotDir, platform,
#                         paste0("accuracyDimensionsTrainObservations-", platform, ".png")))
```

### Table with highest accuracy
```{r eval = FALSE}
tableColumns <- c("algorithm", "classification", "featuresName", "split", "accuracy")
outputParamsExt %>%
  mutate(accuracy = 1 - testError) %>%
  arrange(desc(accuracy)) %>%
  group_by_(.vars = tableColumns) %>%
  select(one_of(tableColumns)) %>%
  filter(row_number() <= 20L) %>%
  htmlTable()
```

### Check for strange accuracy information for logistics regression
```{r eval = FALSE}
tableColumns <- c("algorithm", "classification", "featuresName", "split", "accuracy")
outputParamsExt %>%
  mutate(accuracy = 1 - testError) %>%
  filter(algorithm == "LogReg") %>%
  arrange(desc(accuracy)) %>%
  group_by_(.vars = tableColumns) %>%
  select(one_of(tableColumns)) %>%
  filter(row_number() <= 40L) %>%
  htmlTable()
```

## Visualization of Accuracy/Error rate Random Forest Algorithm
```{r}
applicationsRuntimeExt %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  mutate(accuracy = 1 - testError) %>%
  filter(algorithm == "RF") %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = trainObservations, color = accuracy, 
                            shape = algorithm), alpha = 0.5) +
  labs(color = "Accuracy") +
  ggtitle("Accuracy Dimension/Training Observations (RF)") +
  labs(x = "Dimensions", y = "No of Training Observations") +
  scale_color_gradient(labels = comma, low = "green", high = "red") +
  scale_y_continuous(labels = comma)
# ggsave(file = file.path(plotDir, platform,
#                         paste0("accuracyDimensionsTrainObservations-", platform, ".png")))
```

## Visualization of Accuracy/Error rate Logistics Regression Algorithm
```{r}
applicationsRuntimeExt %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  mutate(accuracy = 1 - testError) %>%
  filter(algorithm == "LogReg") %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = trainObservations, color = accuracy, 
                            shape = algorithm), alpha = 0.5) +
  labs(color = "Accuracy") +
  ggtitle("Accuracy Dimensions/Training Observations (LogReg)") +
  labs(x = "Dimensions", y = "No of Training Observations") +
  scale_color_gradient(labels = comma, low = "green", high = "red") +
  scale_y_continuous(labels = comma)
# ggsave(file = file.path(plotDir, platform,
#                         paste0("accuracyDimensionsTrainObservations-", platform, ".png")))
```

