---
title: "Analysis of Data generate on Spark Cluster Scape01"
author: "Bernhard Bielesch"
output: 
html_document: 
highlight: tango
theme: cerulean
---

```{r setup}
rm(list = ls()) # clean global environment
includeDir <- "../include"
source(file.path(includeDir, "includeRmd.R"))

platform <- "Scape01" # could also be "Local" or "Scape02"
```

## Load necessary data
```{r}
load(file = file.path(dataDir, paste0("outputParamsExt-", platform, ".Rda")))
load(file = file.path(dataDir, paste0("applicationsRuntimeExt-", platform, ".Rda")))
load(file = file.path(dataDir, paste0("jobsRuntime-", platform, ".Rda")))
```

## Calculate overall median, mean and duration
```{r}
summaryAppRuntime <-
  applicationsRuntimeExt %>%
  mutate(coresMax = as.integer(coresMax)) %>%
  group_by(coresMax) %>%
  summarise(medianDuration = median(duration),
            meanDuration = mean(duration),
            sdDuration = sd(duration)) %>%
  arrange(coresMax)

summaryAppRuntime
```
## Build Random Forest to establish Variable Importance

### Preparations for Tree/Forest Building
```{r}
dropCols <- c("id", "appName", "cluster", 
              "startTime", "endTime", 
              "lastUpdated", "features",
              "featuresName", "split", 
              "testObservations", "testError")

factorCols <- c("algorithm", "classification", "coresMax", "execMem")

treeAppRuntime <- 
  applicationsRuntimeExt %>%
  select(-one_of(dropCols)) %>%
  mutate(dimensions = as.integer(dimensions)) %>%
  mutate(trainObservations = as.integer(trainObservations)) %>%
  mutate_at(.cols = factorCols, .fun = funs(factor(.)))

str(treeAppRuntime) # to check all the conversions
Desc(treeAppRuntime)
```
```{r}
# prepare tree building
response <- "duration"
features <- setdiff(names(treeAppRuntime), c(response))
mlformula <- reformulate(termlabels = features, response = response)  
```

### ctree Decision tree
```{r eval = FALSE}
# library(party)
library(partykit)

set.seed(50040)
ctreeModel <- ctree(mlformula, data = treeAppRuntime)
class(ctreeModel)

plot(ctreeModel, type = "simple", main = "Conditional Inference Tree")
```

### Random Forest
```{r}
set.seed(50040)
rfModel <- randomForest(mlformula, data = treeAppRuntime, 
                        ntree = 150, keep.inbag = TRUE, importance = TRUE)
print(rfModel)
importance(rfModel)

varImpPlot(rfModel, main = "Variable Importance for Duration")
```

```{r}
op = par(mfrow=c(1, 2))
plot(rfModel, main = "RF Model")
plot( importance(rfModel), lty=2, pch=16)
lines(importance(rfModel))
```

```{r}
# Partial Dependence Plot
imp = importance(rfModel)
impvar = rownames(imp)[order(imp[, 1], decreasing=TRUE)]
op = par(mfrow=c(3,2))
for (i in seq_along(impvar)) {
  partialPlot(rfModel, treeAppRuntime, impvar[i], xlab=impvar[i],
              main=paste(impvar[i])) # Partial Dependence on...
}
```

### Variable Importance Plot (Alternative)

```{r}
# Extracts variable importance and processed information
varImportance <- data.frame(importance(rfModel)) 
varImportance %<>%
  mutate(variable = as.factor(rownames(.))) %>%
  rename(PercIncMSE = X.IncMSE) %>%
  arrange(desc(PercIncMSE))

varImportance %>%
  ggplot(aes(x = variable, weight = PercIncMSE, fill = variable)) +
  geom_bar() + ggtitle("Variable Importance from Random Forest") + 
  labs(x = "Variable", y = "Variable Importance (Mean Decrease in MSE)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_discrete(limits = as.character(varImportance$variable)) +
  scale_fill_discrete(name = "Variable Name",
                      breaks = as.character(varImportance$variable))
ggsave(file = file.path(plotDir, platform,
                        paste0("varImportance-", platform, ".png")))
```

## Graphical Representations

### This is the distribution of runtimes
```{r}
breaksDurationsLog <- c(250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)
applicationsRuntimeExt %>%
  ggplot(mapping = aes(duration)) + 
  geom_histogram(bins = 40) +
  ggtitle("Distribution of Runtimes") +
  labs(x = "Duration (in ms)", y = "Number of Observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_log10(labels = comma, breaks = breaksDurationsLog)
```
```{r}
breaksDurationsLog <- c(250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)
applicationsRuntimeExt %>%
  ggplot(mapping = aes(duration)) + 
  geom_density() +
  ggtitle("Density of Runtimes") +
  labs(x = "Duration (in ms)", y = "Number of Observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_log10(labels = comma, breaks = breaksDurationsLog)

```

### This is the boxplot of duration across number of cores
```{r}
breaksDurations <- c(250000, 500000, 750000, 1000000, 1250000, 1500000)
applicationsRuntimeExt %>%
  ggplot(mapping = aes(x = as.integer(coresMax),
                       y = duration, group = coresMax)) +
  geom_boxplot() +
  ggtitle("Duration per Number of Cores") + 
  labs(x = "Number of Cores", y = "Duration (in ms)") +
  coord_cartesian(ylim = c(0, 1500000)) + # cut off unnecessary outliers above boxes
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  scale_y_continuous(labels = comma, breaks = breaksDurations)
```

### This is the boxplot of duration across number of cores with notches
```{r}
# Notches are used to compare groups; if the notches of two boxes do not overlap, 
# this suggests that the difference in medians is statistically significantly.
breaksDurations <- c(250000, 500000, 750000, 1000000, 1250000, 1500000)
applicationsRuntimeExt %>%
  ggplot(mapping = aes(x = as.integer(coresMax),
                       y = duration, group = coresMax)) +
  geom_boxplot(notch = TRUE) + # shows confidence interval for each segment, if overlap then no statistical significance 
  ggtitle("Duration per Number of Cores") + 
  labs(x = "Number of Cores", y = "Duration (in ms)") + 
  coord_cartesian(ylim = c(0, 1500000)) + # cut off unnecessary outliers above boxplot
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  scale_y_continuous(labels = comma, breaks = breaksDurations)
```


### Comparison of Mean and Median (huge difference due to non-normal distribution)
```{r}
breaksDurMeanLog <- c(100000, 250000, 500000, 1000000, 1250000)
summaryAppRuntime %>%
  ggplot(mapping = aes(x = as.integer(coresMax), y = meanDuration)) +
  geom_line(color = "blue") + 
  geom_line(color = "green", mapping = aes(y = medianDuration)) +
  ggtitle("Mean & Median Duration per Number of Cores") +
  labs(x = "Number of Cores", y = "Duration (Mean or Median in ms)") + 
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  scale_y_log10(labels = comma, breaks = breaksDurMeanLog) +
  scale_color_continuous(name = "Duration")
```

### Density plot of durations per number of cores
```{r}
require(ggforce)
plot <- 
  applicationsRuntimeExt %>%
  mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
  ggplot(mapping = aes(x = duration)) +
  geom_density() +
  ggtitle("Density Plot of Durations (split by Number of Cores)") +
  labs(x = "Duration (in ms)", y = "Density") +
  scale_x_continuous(labels = comma)

noPages <- n_pages(plot + facet_wrap_paginate(~ coresMax, ncol = 2, nrow = 2))
for (page in 1:noPages) {
  print(plot +  facet_wrap_paginate(~ coresMax, ncol = 2, nrow = 2, page = page))
  # ggsave(file = file.path(plotDir, platform,
  #                         paste0("durationCoresDensity-", platform, ".png")))
}

```

### Density plot of durations per number of cores and dimensions
```{r}

plot <- applicationsRuntimeExt %>%
  mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
  ggplot(mapping = aes(x = duration)) +
  geom_density() +
  ggtitle("Density Plot of Durations (Dimensions and Number of Cores)") +
  labs(x = "Duration", y = "Density") +
  scale_x_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 10))


ncol = 5; nrow = 5
noPages <- n_pages(plot + facet_grid_paginate(dimensions ~ coresMax, 
                                              ncol = ncol, nrow = nrow))
for (page in 1:noPages) {
  print(plot +  facet_grid_paginate(dimensions ~ coresMax, 
                                    ncol = ncol, nrow = nrow, page = page))
}
```

### Density plot of durations per number of cores
```{r}
breaksDurationsLog <- c(250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)
applicationsRuntimeExt %>%
  mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
  ggplot(mapping = aes(x = dimensions)) +
  geom_jitter(mapping = aes(y = duration, color = coresMax)) +
  labs(x = "Dimensions", y = "Duration (in ms)") + 
  scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
  # facet_wrap(~ coresMax, ncol = 1) + 
  ggtitle("Duration per Dimension/Cores")
ggsave(file = file.path(plotDir, platform,
                        paste0("durationDimensionsCores-", platform, ".png")))
```

## Get rid of outlier that seems wrong
```{r}
tableColumns <- c("id", "appName", "duration")
applicationsRuntimeExt %>%
  filter(dimensions == "1176") %>%
  arrange(duration) %>%
  select(one_of(tableColumns)) %>%
  head(5) %>%
  htmlTable()
```
```{r}
applicationsRuntimeExt %<>% # change underlying data
  filter(id != "app-20170622153528-2249")
```

## Try previous plot again
```{r}
breaksDurationsLog <- c(250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)
applicationsRuntimeExt %>%
  mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
  ggplot(mapping = aes(x = dimensions)) +
  geom_jitter(mapping = aes(y = duration, color = coresMax)) +
  labs(x = "Dimensions", y = "Duration (in ms)") + 
  scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
  # facet_wrap(~ coresMax, ncol = 1) + 
  ggtitle("Duration per Dimension/Cores")
ggsave(file = file.path(plotDir, platform,
                        paste0("durationDimensionsCores-", platform, ".png")))
```

## Mean and sd duration per memory size
```{r}
summaryAppMemRuntime <-
  applicationsRuntimeExt %>%
  mutate(coresMax = as.integer(coresMax)) %>%
  group_by(execMem, coresMax) %>%
  summarise(meanDuration = mean(duration),
            medianDuration = median(duration),
            sdDuration = sd(duration)) %>%
  setorder(execMem, coresMax)

summaryAppMemRuntime
```

## Graphical presentation

### Median Duration per Number of Cores and Memory
```{r}
breaksDurMeanLog <- c(150000, 200000, 250000, 350000, 450000)
summaryAppMemRuntime %>%
  ggplot(mapping = aes(x = coresMax, y = medianDuration, group = execMem, 
                       col = execMem, shape = execMem)) +
  geom_line() + geom_point() +
  ggtitle("Median Duration per Number of Cores") +
  labs(x = "Number of Cores", y = "Median Duration (in ms)") + 
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  scale_y_log10(labels = comma, breaks = breaksDurMeanLog) +
  scale_color_discrete(name = "Memory") +
  scale_shape_discrete(name = "Memory") + 
  theme(legend.position = "right")
```

```{r}
breaksDurationsLog <- c(250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)
execMems <- c("1G", "2G", "3G")

for (execMem in execMems) {
  plot <- applicationsRuntimeExt %>%
    filter(execMem == execMem) %>%
    mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
    mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
    ggplot(mapping = aes(x = dimensions)) +
    geom_jitter(mapping = aes(y = duration, color = coresMax)) +
    labs(x = "Dimensions", y = "Duration (in ms)") + 
    scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
    ggtitle(paste0("Duration per Dimensions/Cores: Execution Memory ", execMem))
  print(plot)
  ggsave(file = file.path(plotDir, platform,
                          paste0("durationDimensionsCores-", platform, "-ExecMem", execMem, ".png")))
}

```

### Duration facetted per dimensions and cores
```{r}
dimensions <- as.character(sort(as.integer(unique(applicationsRuntimeExt[["dimensions"]]))))
breaksDurationsLog <- c(65000, 75000, 125000, 200000, 250000, 500000, 1000000, 2000000, 3000000, 4000000, 5000000, 6000000)

for (dimension in dimensions) {
  
  plot <- applicationsRuntimeExt %>%
    filter(dimensions == dimension) %>%
    mutate(coresMax = factor(coresMax, levels = sort(as.integer(unique(coresMax))))) %>%
    mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
    ggplot(mapping = aes(x = coresMax, y = duration)) +
    geom_boxplot(notch = TRUE) +
    ggtitle(paste0("Durations per number of cores (Dimensions: ", dimension, ")")) +
    labs(x = "maxCores", y = "Duration (in ms)") +
    scale_y_log10(label = comma, breaks = breaksDurationsLog) +
    theme(plot.title = element_text(size = 10))
  print(plot)
  ggsave(file = file.path(plotDir, platform,
                          paste0("durationDimensions-", platform, "-Dim", dimension, ".png")))
}
```

## Median Duration per cores for different dimensions
```{r}
summaryAppDimRuntime <-
  applicationsRuntimeExt %>%
  mutate(coresMax = as.integer(coresMax)) %>%
  group_by(coresMax, dimensions) %>%
  summarise(meanDuration = mean(duration),
            medianDuration = median(duration),
            sdDuration = sd(duration)) %>%
  setorder(dimensions, coresMax)

summaryAppDimRuntime
```

```{r}
breaksDurMeanLog <- c(150000, 300000, 600000, 1500000, 3500000)
summaryAppDimRuntime %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions)), 
                                                       decreasing = TRUE))) %>% 
  ggplot(mapping = aes(x = coresMax, y = medianDuration, group = dimensions, 
                       col = dimensions)) +
  geom_line() + geom_point() +
  ggtitle("Median Duration per Dimensions/Number of Cores") +
  labs(x = "Number of Cores", y = "Median Duration (in ms)") + 
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  scale_y_log10(labels = comma, breaks = breaksDurMeanLog) +
  theme(legend.position = "right")

ggsave(file = file.path(plotDir, platform,
                        paste0("medianDurationSpeedup-", platform, ".png")))
```

## SpeedUp 
```{r}
summaryAppSpeedup <-
  summaryAppDimRuntime %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>% 
  group_by(dimensions) %>%
  setorder(dimensions, coresMax) %>%
  mutate(speedup = lag(medianDuration)/medianDuration - 1) %>%
  mutate(speedup = ifelse(is.na(speedup), 0, speedup)) %>%
  mutate(speedupCum = cumsum(speedup)) # cumulative sum of speedUp
summaryAppSpeedup
```

```{r}
summaryAppSpeedup %>%
  ggplot(mapping = aes(x = coresMax, y = speedup, group = dimensions, 
                       fill = dimensions)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  ggtitle("SpeedUp per Number of Cores/Dimensions") +
  labs(x = "Number of Cores", y = "Speedup") + 
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2), limits = c(1.5,15.5)) +
  theme(legend.position = "right")

ggsave(file = file.path(plotDir, platform,
                        paste0("speedupFactorDimCores-", platform, ".png")))
```

```{r}
summaryAppSpeedup %>%
  na.omit() %>%
  setorder(-speedup) %>%
  head(5)
```

```{r}
summaryAppSpeedup %>%
  na.omit() %>%
  setorder(speedup) %>%
  head(5)
```

```{r}
breaksSpeedupCum <- c(1,2,3,4,5,6,7,8,9,10)
summaryAppSpeedup %>%
  ungroup(dimensions) %>%  
  mutate(dimensions = factor(dimensions, levels = rev(levels(dimensions)))) %>%
  group_by(dimensions)  %>%
  mutate(speedup = ifelse(speedup == 0, 0, speedup - 1)) %>%
  mutate(speedupCum = cumsum(speedup)) %>%

  ggplot(mapping = aes(x = coresMax, y = speedupCum, group = dimensions, 
                       col = dimensions)) +
  geom_line() + geom_point() +
  ggtitle("Cumulative SpeedUp per Number of Cores/Dimensions") +
  labs(x = "Number of Cores", y = "Cumulative Speedup") + 
  scale_x_continuous(breaks = seq(from = 2, to = 14, by = 2)) +
  # scale_y_log10(breaks = breaksSpeedupCum) +
  theme(legend.position = "right")

ggsave(file = file.path(plotDir, platform,
                        paste0("speedupCumulativeDimCores-", platform, ".png")))
```

