---
title: "Analysis of Data generate on Single-node Spark Cluster"
author: "Bernhard Bielesch"
output: 
  html_document: 
    highlight: tango
    theme: cerulean
---
```{r setup}
rm(list = ls()) # clean global environment
includeDir <- "../include"
source(file.path(includeDir, "includeRmd.R"))

platform <- "Local" # could also be "Local" or "Scape02"
```

## Load necessary data
```{r}
#####################################################################

load(file = file.path(dataDir, paste0("outputParamsExt-", platform, ".Rda")))
load(file = file.path(dataDir, paste0("phasesRuntime-", platform, ".Rda")))
```

## Analze max, min, mean, median and sd of data
```{r}
phasesRuntime %>%
  group_by_(.dots = c("phase")) %>%
  summarize(maxDuration = max(phaseDuration),
            minDuration = min(phaseDuration),
            meanDuration = mean(phaseDuration),
            medianDuration = median(phaseDuration),
            sdDuration = sd(phaseDuration)) %>%
  mutate(phase = as.character(phase))
```

## Analyze max, min, mean, median and sd of percentages
```{r}

roundCols <- c("maxPerc", "minPerc", "meanPerc", "medianPerc", "sdPerc")
phasesRuntime %>%
  group_by_(.dots = c("phase")) %>%
  summarize(maxPerc = max(phasePerc),
            minPerc = min(phasePerc),
            meanPerc = mean(phasePerc),
            medianPerc = median(phasePerc),
            sdPerc = sd(phasePerc)) %>%
  mutate_at(.vars = roundCols, .funs = funs(round(., 3))) %>%
  mutate(phase = as.character(phase))
```

## Variable Importance for modelBuilding

### Build Random Forest Model
```{r}
dropCols <- c("id", "appName", "cluster", "features", "featuresName", "split", 
              "appDuration", "phasePerc", "duration", "testObservations", "testError")
factorCols <- c("algorithm", "classification")
integerCols <- c("dimensions", "trainObservations")

treePhaseLearning <- 
  phasesRuntime %>%
  ungroup() %>%
  filter(phase == "learning") %>%
  select(-one_of(c("phase"))) %>%
  select(-one_of(dropCols)) %>%
  mutate_at(.vars = integerCols, .funs = funs(as.integer(.))) %>%
  mutate_at(.vars = factorCols, .fun = funs(factor(.)))

Desc(treePhaseLearning)
```

```{r}
# prepare tree building
response <- "phaseDuration"
features <- setdiff(names(treePhaseLearning), c(response))
mlformula <- reformulate(termlabels = features, response = response)  

set.seed(50040)
rfPhaseLearning <- randomForest(mlformula, data = treePhaseLearning, 
                        ntree = 150, keep.inbag = TRUE, importance = TRUE)
importance(rfPhaseLearning)
```

### Extracts variable importance and processed information
```{r}
varImportance <- data.frame(importance(rfPhaseLearning)) 
varImportance %<>%
  mutate(variable = as.factor(rownames(.))) %>%
  rename(PercIncMSE = X.IncMSE) %>%
  arrange(desc(PercIncMSE))
               
varImportance %>%
  ggplot(aes(x = variable, weight = PercIncMSE, fill = variable)) +
  geom_bar() + ggtitle("Learning Phase: Variable Importance from Random Forest") + 
  xlab("Variable") + ylab("Variable Importance (Mean Decrease in MSE)") +
  scale_x_discrete(limits = as.character(varImportance$variable)) +
  scale_fill_discrete(name = "Variable Name",
                      breaks = as.character(varImportance$variable))
```

### Curve of Learning phase duration vs. dimensions and trainObservations
```{r}
breaksDurationsLog <- c(50000, 250000, 500000, 750000, 1000000, 1250000)
  
phasesRuntime %>%
  ungroup() %>%
  filter(phase == "learning") %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = phaseDuration, color = trainObservations), alpha = 0.5) +
  scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
  scale_colour_continuous(labels = comma) +
  labs(color = "Training Observations") +
  ggtitle("Learning Phase: Duration per dimensions") +
  labs(x = "Dimensions", y = "Duration (ms)")

ggsave(file = file.path(plotDir, platform, 
                        paste0("phaseDurationLearning-", platform, ".png")))
```

## Variable Importance for Data Phase

### Build Random Forest Model
```{r}
dropCols <- c("id", "appName", "cluster", "features", "featuresName", "split", 
              "appDuration", "phasePerc", "duration", "testObservations", "testError")
factorCols <- c("algorithm", "classification")
integerCols <- c("dimensions", "trainObservations")

treePhaseData <- 
  phasesRuntime %>%
  ungroup() %>%
  filter(phase == "data") %>%
  select(-one_of(c("phase"))) %>%
  select(-one_of(dropCols)) %>%
  mutate_at(.vars = integerCols, .funs = funs(as.integer(.))) %>%
  mutate_at(.vars = factorCols, .fun = funs(factor(.)))

Desc(treePhaseData)
```

```{r}
# prepare tree building
response <- "phaseDuration"
features <- setdiff(names(treePhaseData), c(response))
mlformula <- reformulate(termlabels = features, response = response)  

set.seed(50040)
rfPhaseData <- randomForest(mlformula, data = treePhaseData, 
                        ntree = 150, keep.inbag = TRUE, importance = TRUE)
importance(rfPhaseData)
```

### Extracts variable importance and processed information
```{r}
varImportance <- data.frame(importance(rfPhaseData)) 
varImportance %<>%
  mutate(variable = as.factor(rownames(.))) %>%
  rename(PercIncMSE = X.IncMSE) %>%
  arrange(desc(PercIncMSE))
               
varImportance %>%
  ggplot(aes(x = variable, weight = PercIncMSE, fill = variable)) +
  geom_bar() + ggtitle("Data Phase: Variable Importance from Random Forest") + 
  xlab("Variable") + ylab("Variable Importance (Mean Decrease in MSE)") +
  scale_x_discrete(limits = as.character(varImportance$variable)) +
  scale_fill_discrete(name = "Variable Name",
                      breaks = as.character(varImportance$variable))
```

### Curve of Data phase duration vs. dimensions and trainObservations
```{r}
breaksDurationsLog <- c(25000, 50000, 100000, 150000, 250000)
phasesRuntime %>%
  ungroup() %>%
  filter(phase == "data") %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = phaseDuration, color = trainObservations), alpha = 0.5) +
  scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
  scale_colour_continuous(labels = comma) +
  labs(color = "Training Observations") +
  ggtitle("Data Phase: Duration per dimensions") +
  labs(x = "Dimensions", y = "Duration (ms)")
ggsave(file = file.path(plotDir, platform, 
                        paste0("phaseDurationData-", platform, ".png")))
```
## Variable Importance for Prediction Phase

### Build Random Forest Model
```{r}
dropCols <- c("id", "appName", "cluster", "features", "featuresName", "split", 
              "appDuration", "phasePerc", "duration", "trainObservations", "testError")
factorCols <- c("algorithm", "classification")
integerCols <- c("dimensions", "testObservations")

treePhasePrediction <- 
  phasesRuntime %>%
  ungroup() %>%
  filter(phase == "prediction") %>%
  select(-one_of(c("phase"))) %>%
  select(-one_of(dropCols)) %>%
  mutate_at(.vars = integerCols, .funs = funs(as.integer(.))) %>%
  mutate_at(.vars = factorCols, .fun = funs(factor(.)))

Desc(treePhasePrediction)
```


```{r}
# prepare tree building
response <- "phaseDuration"
features <- setdiff(names(treePhasePrediction), c(response))
mlformula <- reformulate(termlabels = features, response = response)  

set.seed(50040)
rfPhasePrediction <- randomForest(mlformula, data = treePhasePrediction, 
                        ntree = 150, keep.inbag = TRUE, importance = TRUE)
importance(rfPhasePrediction)
```

### Extracts variable importance and processed information
```{r}
varImportance <- data.frame(importance(rfPhasePrediction)) 
varImportance %<>%
  mutate(variable = as.factor(rownames(.))) %>%
  rename(PercIncMSE = X.IncMSE) %>%
  arrange(desc(PercIncMSE))
               
varImportance %>%
  ggplot(aes(x = variable, weight = PercIncMSE, fill = variable)) +
  geom_bar() + ggtitle("Prediction Phase: Variable Importance from Random Forest") + 
  xlab("Variable") + ylab("Variable Importance (Mean Decrease in MSE)") +
  scale_x_discrete(limits = as.character(varImportance$variable)) +
  scale_fill_discrete(name = "Variable Name",
                      breaks = as.character(varImportance$variable))
```

### Curve of Prediction phase duration vs. dimensions and testObservations
```{r}
breaksDurationsLog <- c(25000, 50000, 100000, 150000, 250000, 350000)
phasesRuntime %>%
  ungroup() %>%
  filter(phase == "prediction") %>%
  mutate(dimensions = factor(dimensions, levels = sort(as.integer(unique(dimensions))))) %>%
  ggplot(mapping = aes(x = dimensions)) + 
  geom_jitter(mapping = aes(y = phaseDuration, color = testObservations), alpha = 0.5) +
  scale_y_log10(labels = comma, breaks = breaksDurationsLog) +
  scale_colour_continuous(labels = comma) +
  labs(color = "Test Observations") +
  ggtitle("Prediction Phase: Duration per dimensions") +
  labs(x = "Dimensions", y = "Duration (ms)")
ggsave(file = file.path(plotDir, platform, 
                        paste0("phaseDurationPrediction-", platform, ".png")))
```